### 2014-10-20 Publications/2013-TMI/03-introduction.tex
update TMI paper

\section{Introduction}
\label{sec:introduction}

\IEEEPARstart{D}{iffusion} \gls{mri} is a widely used family
of \gls{mri} techniques \citep{sundgren_diffusion_2004} which 
accounts for a growing interest in its application to structural 
connectivity analyses of the brain at the macro-scale
\citep{craddock_imaging_2013}. This emerging field exploits
\gls{dwi} data to derive the local axonal structure at each imaging voxel 
\citep{basser_microstructural_1996} and estimate a whole-brain mapping of fiber 
tracts \citep{hagmann_mr_2012,johansen-berg_using_2009} represented by 
trajectories reconstructed from the local information.
This comprehensive map of neural connections of the brain is called the 
\emph{connectome} \citep{hagmann_diffusion_2005,sporns_human_2005}. The connectome
analysis is a promising tool for neuroscience \citep{morgan_why_2013} and 
for the clinical application \citep{griffa_structural_2013} as well.

% REMOVE paragraph (Dom, 1st comment)
%Early \gls{dwi} research mainly focused on the improvements of imaging 
%methodologies by a better understanding of the diffusion effect, and improving
%image reconstruction methodologies. Currently, the connectome extraction 
%and analysis relies on a large chain of sophisticated computational techniques
%including acquisition, reconstruction, modeling and model fitting, image segmentation
%and registration, fibre tracking, connectivity mapping, visualization, statistical 
%group studies, and inference \citep{daducci_connectome_2012}. This growing complexity 
%has given rise to challenging issues towards reliable structural information about 
%the neuronal tracts presented in, and also difficulties on the statistical analysis 
%\citep{bullmore_complex_2009,meskaldji_comparing_2013}. To our knowledge, studies 
%regarding the impact of the different sources of inaccuracy and variability along
%the toolchain for connectivity analysis \citep{jones_twenty-five_2010,
%cheng_characteristics_2012,jones_white_2013} are scarce.

\subsection{\Gls{dwi} data overview}
\label{sec:dwi_overview}
\gls{dwi} data are usually acquired with \gls{epi} schemes
as they allow for very fast acquisition of large sets of slices each
acquired after a single excitation \citep{mukherjee_diffusion_2008,
soares_hitchhikers_2013}. \Gls{epi} volumes are collected
in sequence and each one represents one probed diffusion direction.
%This first acquisition is called \emph{raw} data or \emph{diffusion
%signal}. One or more volumes without directional encoding (so-called \emph{baseline},
%$b = 0$, \emph{low $b$}, or just \emph{b0} volumes) are interleaved as
%reference, at the least one is acquired at the beginning of the
%\gls{epi} dataset. These B0 are acquired without direction gradient,
%showing a T2-like contrast.

\Gls{epi} are known to suffer from geometrical distortions and artifacts
due to, mainly, three sources: the subject motion in between acquiring 
different sampling directions, the induced \emph{eddy currents} on the scanner 
coils due to the gradient switching \citep{obrien_3d_2013}, and, finally, 
distortions caused by the magnetic susceptibility inhomogeneity
present at tissue interface. \emph{Susceptibility distortions} happen along the 
phase-encoding direction, and are most appreciable in the frontal part of 
the brain due to the air/tissue interface located around the sinuses.
Two implications are associated to this artifact: the signal loss caused on
highly distorted regions, and a significant added difficulty in 
registering with structural images (e.g. T1-weighted).

After correcting for artifacts, the difficulty of performing analyses directly
on the \emph{directionally dependent} raw data is not affordable. Thus, 
\gls{dwi} data is reconstructed using one of a numerous set of readily 
available algorithms. The outcome is a voxel-wise model of diffusion shape 
(\citep{daducci_quantitative_2013} contains a comprehensive review of 
state-of-art methodologies). On top of the
chosen model, some scalar maps describing tissue features are generally
computed (e.g. \gls*{fa}, and \gls*{adc}).
%For instance, the most popular \gls{dwi}-derived scalar map 
%is \gls{fa} which depicts the isotropy of water diffusion inside the brain.
%Another widely studied map is the \gls{adc} (or \gls{md} that has a dual definition),
%which reflects the intensity of the diffusion effect. Many other
%directionally invariant measurements have been defined. For a comprehensive study, 
%we refer to \citep{ennis_orthogonal_2006}.

@@

% the low resolution achievable with \gls{dwi} 
%(typically around ). An additional complication specific to
%\gls{dwi} is the \gls{csf} \emph{contamination} \citep{metzler-baddeley_how_2012},
%that is a particular \gls{pve} in which the signal sensed inside the affected voxel is 
%linear with respect the \gls{gm} and \gls{wm} contributions, but non-linear with
%respect \gls{csf}.

@@

One approach to correcting the susceptibility distortions was proposed
with the earliest \gls{epi} sequences used in functional \gls{mri}, and
relies on the acquisition of extra \gls{mri} data. Generally, a \gls{gre} is used to 
obtain a map of magnitude and phase of the actual magnetic field inside the
scanner. Based on this \emph{fieldmap} and the theory underlying the 
distortion, \citep{jezzard_correction_1995} proposed a correction 
methodology. A number of forked and improved methodologies have been 
developed to correct for the susceptibility distortion and received the
generic name of ``fieldmap correction'' techniques \cite{hsu_correction_2009,
reber_correction_1998}.

Some other retrospective methodologies do not make use of the fieldmaps, 
as explicitly modeling the distortion \citep{andersson_modeling_2001},
registering with (anatomically correct) T2-weighted \gls{mri} 
\citep{kybic_unwarping_2000,studholme_accurate_2000,li_accounting_2007,
tao_variational_2009}, or acquiring an 
extra B0 image with reversed phase encoding direction 
\citep{andersson_how_2003,holland_efficient_2010}.

@@

A precise delineation of the \gls{csf}, \gls{gm} and \gls{wm} interfacing surfaces
is required with sub-pixel resolution, because whereas tractography algorithms are
very good at estimating the location of bundles in deep \gls{wm}, they are not
god (yet) at identifying where the tracts project into the \gls{gm} 
(limitations on the termination criteria, \citep{jbabdi_tractography:_2011,
craddock_imaging_2013}).
Thus, a \gls{wm} segmentation is required by most of the available tractography 
algorithms to \emph{filter} the resulting tractogram. The \gls{gm}-\gls{wm}
interface is necessary to locate the starting and ending points of the detected
fiber bundles, and \gls{csf}-\gls{wm} surface is critical for pruning spurious 
and discontinued fiber bundles.

@@

None of the presented methods have claimed for definite results, mainly due to the lack of 
a \emph{gold-standard} evaluation methodology. Most of them are tested only on certain
regions, or do not provide sub-pixel resolution results
\citep{hadjiprocopis_unbiased_2005,liu_brain_2007,awate_multivariate_2008,lu_segmentation_2008,
han_experimental_2009}. Generally, results obtained with high resolution atlas co-registration
(\autoref{sec:registration}) are more compelling, limiting the activity on \gls{dwi} 
segmentation. However, recently \citep{kumazawa_improvement_2013} proposed to include this
task within reconstruction algorithms. Finally, some \emph{golden}-standard evaluation
frameworks for task-based validation have been proposed for the segmentation results
(e.g. lesion detection in visceral organs \citep{jha_task-based_2012}).
\todo[inline]{reference to a summary table}

@@

The significant benefits of exploiting the anatomical \gls{mri} when 
segmenting the \gls{dwi} data have been demonstrated \cite{zollei_improved_2010}, 
justifying the use of the shape prior information. To our knowledge, there is
no study simultaneously taking advantage from segmentation or distortion correction
tasks to be applied to the co-registration problem.

@@

All the available \gls{dwi}-derived models and scalar maps 
are very sensitive to distortions, \gls{pve}, noise and axonal dispersion 
\citep{jbabdi_tractography:_2011}, what yields numerous challenges in the
processing tasks (artifact correction, segmentation, registration and 
parcellation). Finally, as a result a vast number of challenges and pitfalls 
emerge regarding accuracy (correctness) and precision (reproducibility) of 
tractography-based analyses \citep{jones_white_2013} and connectome
analyses.

>>>>

% -*- root: 00-main.tex -*-
\section{Introduction}
\label{sec:introduction}

\IEEEPARstart{D}{iffusion} \gls{mri} is a widely used family
  of \gls{mri} techniques \citep{sundgren_diffusion_2004} which
  accounts for a growing interest in its application to structural
  connectivity analyses of the brain at the macro-scale
  \citep{craddock_imaging_2013}.
This emerging field exploits \gls{dwi} data to derive the local axonal structure at
  each imaging voxel \citep{basser_microstructural_1996} and estimate a whole-brain
  mapping of fiber tracts \citep{hagmann_mr_2010,berg_using_2009} trajectories
  reconstructed from the local information.
This comprehensive map of neural connections of the brain is called the
  \emph{connectome} \citep{hagmann_from_2005,sporns_human_2005}.
The connectome analysis is a promising tool for neuroscience \citep{morgan_why_2013}
  and for the clinical application \citep{griffa_structural_2013} as well.

% REMOVE paragraph (Dom, 1st comment)
%Early \gls{dwi} research mainly focused on the improvements of imaging
%methodologies by a better understanding of the diffusion effect, and improving
%image reconstruction methodologies.
%Currently, the connectome extraction
%and analysis relies on a large chain of sophisticated computational techniques
%including acquisition, reconstruction, modeling and model fitting, image segmentation
%and registration, fibre tracking, connectivity mapping, visualization, statistical
%group studies, and inference \citep{daducci_connectome_2012}.
%This growing complexity
%has given rise to challenging issues towards reliable structural information about
%the neuronal tracts presented in, and also difficulties on the statistical analysis
%\citep{bullmore_complex_2009,meskaldji_comparing_2013}.
%To our knowledge, studies
%regarding the impact of the different sources of inaccuracy and variability along
%the toolchain for connectivity analysis \citep{jones_twenty-five_2010,
%cheng_characteristics_2012,jones_white_2013} are scarce.

\subsection{\Gls{dwi} data overview}
\label{sec:dwi_overview}
\gls{dwi} data are usually acquired with \gls{epi} schemes as they allow for very
  fast acquisition of large sets of slices each acquired after a single excitation
  \citep{mukherjee_diffusion_2008,soares_hitchhikers_2013}.
\Gls{epi} volumes are collected in sequence and each one represents one probed
  diffusion direction.
%This first acquisition is called \emph{raw} data or \emph{diffusion
%signal}.
%One or more volumes without directional encoding (so-called \emph{baseline},
%$b = 0$, \emph{low $b$}, or just \emph{b0} volumes) are interleaved as
%reference, at the least one is acquired at the beginning of the
%\gls{epi} dataset.
%These B0 are acquired without direction gradient,
%showing a T2-like contrast.

\Gls{epi} are known to suffer from geometrical distortions and artifacts
  due to, mainly, three sources: the subject motion in between acquiring
  different sampling directions, the induced \emph{eddy currents} on the scanner
  coils due to the gradient switching \citep{brien_3d_2013}, and, finally,
  distortions caused by the magnetic susceptibility inhomogeneity
  present at tissue interfaces.
\emph{Susceptibility distortions} happen along the phase-encoding direction,
  and are most appreciable in the frontal part of the brain due to the air/tissue
  interface located around the sinuses.
Two implications are associated to this artifact: the signal loss caused on
  highly distorted regions, and a significant added difficulty in
  registering with structural images (e.g.
T1-weighted).

After correcting for artifacts, the difficulty of performing analyses directly
  on the \emph{directionally dependent} raw data is not affordable.
Thus, \gls{dwi} data is reconstructed using one of a numerous set of readily
  available algorithms.
The outcome is a voxel-wise model of diffusion shape (\citep{daducci_quantitative_2014}
  contains a comprehensive review of state-of-art methodologies).
On top of the chosen model, some scalar maps describing tissue features are generally
  computed (e.g.
\gls*{fa}, and \gls*{adc}).
%For instance, the most popular \gls{dwi}-derived scalar map
%is \gls{fa} which depicts the isotropy of water diffusion inside the brain.
%Another widely studied map is the \gls{adc} (or \gls{md} that has a dual definition),
%which reflects the intensity of the diffusion effect.
%Many other
%directionally invariant measurements have been defined.
%For a comprehensive study,
%we refer to \citep{ennis_orthogonal_2006}.

@@

% the low resolution achievable with \gls{dwi}
%(typically around ).
%An additional complication specific to
%\gls{dwi} is the \gls{csf} \emph{contamination} \citep{metzler-baddeley_how_2012},
%that is a particular \gls{pve} in which the signal sensed inside the affected voxel is
%linear with respect the \gls{gm} and \gls{wm} contributions, but non-linear with
%respect \gls{csf}.

@@

One approach to correcting the susceptibility distortions was proposed with the
  earliest \gls{epi} sequences used in functional \gls{mri}, and relies on the
  acquisition of extra \gls{mri} data.
Generally, a \gls{gre} is used to obtain a map of magnitude and phase of the actual
  magnetic field inside the scanner.
Based on this \emph{fieldmap} and the theory underlying the distortion,
  \citep{jezzard_correction_1995} proposed a correction methodology.
A number of forked and improved methodologies have been developed to correct
  for the susceptibility distortion and received the generic name of
  ``fieldmap correction'' techniques \cite{hsu_correction_2009,reber_correction_1998}.

Some other retrospective methodologies do not make use of the fieldmaps, as explicitly
  modeling the distortion \citep{andersson_modeling_2001}, registering with (anatomically
  correct) T2-weighted \gls{mri} \citep{kybic_unwarping_2000,studholme_accurate_2000,
  li_accounting_2007,tao_variational_2009}, or acquiring an extra B0 image with reversed
  phase encoding direction \citep{andersson_how_2003,holland_efficient_2010}.

@@

A precise delineation of the \gls{csf}, \gls{gm} and \gls{wm} interfacing surfaces
  is required with sub-pixel resolution, because whereas tractography algorithms are
  very good at estimating the location of bundles in deep \gls{wm}, they are not
  god (yet) at identifying where the tracts project into the \gls{gm}
  (limitations on the termination criteria, \citep{jbabdi_tractography_2011,
  craddock_imaging_2013}).
Thus, a \gls{wm} segmentation is required by most of the available tractography
algorithms to \emph{filter} the resulting tractogram.
The \gls{gm}-\gls{wm} interface is necessary to locate the starting and ending points
  of the detected fiber bundles, and \gls{csf}-\gls{wm} surface is critical for pruning
  spurious and discontinued fiber bundles.

@@

None of the presented methods have claimed for definite results, mainly due to the lack of
  a \emph{gold-standard} evaluation methodology.
Most of them are tested only on certain
  regions, or do not provide sub-pixel resolution results
  \citep{hadjiprocopis_unbiased_2005,liu_brain_2007,awate_multivariate_2008,lu_segmentation_2008,
  han_experimental_2009}.
Generally, results obtained with high resolution atlas co-registration
  (\autoref{sec:registration}) are more compelling, limiting the activity on \gls{dwi}
  segmentation.
However, recently \citep{kumazawa_improvement_2013} proposed to include this
  task within reconstruction algorithms.
Finally, some \emph{golden}-standard evaluation frameworks for task-based validation have been
  proposed for the segmentation results (e.g. lesion detection in visceral organs
  \citep{jha_taskbased_2012}).
\todo[inline]{reference to a summary table}

@@

The significant benefits of exploiting the anatomical \gls{mri} when
segmenting the \gls{dwi} data have been demonstrated \cite{zoellei_improved_2010},
justifying the use of the shape prior information.
To our knowledge, there is
no study simultaneously taking advantage from segmentation or distortion correction
tasks to be applied to the co-registration problem.

@@

All the available \gls{dwi}-derived models and scalar maps
are very sensitive to distortions, \gls{pve}, noise and axonal dispersion
\citep{jbabdi_tractography_2011}, what yields numerous challenges in the
processing tasks (artifact correction, segmentation, registration and
parcellation).
Finally, as a result a vast number of challenges and pitfalls
emerge regarding accuracy (correctness) and precision (reproducibility) of
tractography-based analyses \citep{jones_white_2013} and connectome
analyses.

### 2014-10-20 Publications/2013-TMI/04-methods.tex
update TMI paper

1\section{Methods}
\label{sec:methods}
%
\subsection{Related work}
\label{sec:methods_background}

We suggest clustering the 
current methodologies of template-based segmentation methods into three groups. 
The first group typically adds a shape prior term to the energy functional of 
an evolving active contour \citep{bresson_variational_2006,
chan_level_2005,chen_using_2002,cremers_kernel_2006,gastaud_combining_2004}.
These methods generally have an explicit description of the expected relative boundary 
locations of the object to be delineated, and some even model the statistical deviations
from this average shape. By including a coordinates mapping, it is possible to perform
active contours based registration between timesteps in a time-series or between different
images \citep{bertalmio_morphing_2000,wyatt_map_2003,paragios_level_2003,vemuri_joint_2003,
yezzi_variational_2003}.
A summary of these second set of techniques is performed in \citep{droske_mumfordshah_2009},
proposing two different approaches to applying the Mumford-Shah \citep{mumford_optimal_1989}
functional in joint registration and segmentation. Finally, a derivation of the 
latter groups is composed by atlas-based segmentation methods 
\citep{gorthi_segmentation_2009,gorthi_active_2011,pohl_unifying_2005,
pohl_bayesian_2006,wang_joint_2006}, where the prior 
imposes consistent voxel-based classification of contiguous regions.
A comprehensive summary of the \gls{acwe}-derived methodologies with special attention
to joint registration-segmentation methods is found in \citep{gorthi_active_2011}.
\Gls{acwe} and level-set based formulations have been widely applied in image
processing, as reviewed in \citep{suri_shape_2002}.

@@

Even though the historical evolution
of joint segmentation and registration procedures is strongly linked
to the level sets approximation, our work is based in derivations of
the \gls{acwe} framework \citep{chan_active_2001}, and related to
\citep{le_guyader_combined_2011}.
{\color{red} {The main diverging points with respect to
\citep{gorthi_active_2011} are: 1) there is no need for an explicit level set function
$\Phi_G$, as we replace the level set gradient computation $N_{\Phi_G}$ with shape
gradients \citep{jehan-besson_dream2s:_2003,herbulot_segmentation_2006};
2) regularization is also based on linear diffusion smoothing \citep{thirion_image_1998},
but we replace the Gaussian filtering by other constraints
studied in \citep{nagel_investigation_1986} to better the problem;
3) optimization is applied in the spectral
domain, observing anisotropic and inhomogeneous mappings along each direction.
With respect \citep{le_guyader_combined_2011}, the main differences are
the distance function, and the spectral solution to the optimization updates,
as we shall cover in \autoref{sec:numerical_implementation}.}}

@@

{\color{red} {Using the region descriptors derived in \autoref{sec:methods_map}, we propose
an \gls{adf}-like, \gls{acwe}-based, piece-wise constant, image segmentation
model (where the unknown is the deformation field)
\cite{chan_active_2001} with the energy functional obtained in 
\eqref{eq:final_map_energy}. This inverse problem is ill-posed
\cite{bertero_ill-posed_1988,hadamard_sur_1902}.
In order to account for deformation field regularity and to render the 
problem well-posed, we include limiting and regularization terms into 
the energy functional \cite{morozov_linear_1975,tichonov_solution_1963}:
%
\begin{multline}
E(u) = \sum\limits_l \int_{\Omega'_l} \mdist{f}{l} \,d\vec{x} \\
+ \int \vec{u}^T \, A \, \vec{u} \, d\vec{x} + \int \tr\{(\nabla \vec{u}^T)^T B (\nabla \vec{u}^T)\} d\vec{x}
\label{eq:complete_energy}
\end{multline}
%
These regularity terms ensure that the segmenting contours in 
\gls{dwi} space are still close to their native shape. The model
easily allows to incorporate inhomogeneous and anisotropic 
regularization \cite{nagel_investigation_1986} to better regularize
the \gls{epi} distortion.}}

@@

\subsubsection{Deformation field}
\label{sec:deformation_field}
The transformation from ``T1-'' into ``\gls{dwi}-'' coordinate space is 
achieved through a dense deformation field $u(\vec{x})$, such that:
\begin{equation}
\vec{c}_i' = U\{\vec{c}_i\} = \vec{c}_i + u(\vec{c}_i),
\label{eq:nodes_tfm}
\end{equation}
where $U$ is defined in \eqref{eq:deformation_field}. Since the nodes of the anatomical 
surfaces might lay off-grid, it is required to derive $u(\vec{x})$ from a discrete 
set of parameters $\{\vec{u}_k\}_{k=1 \ldots K}$. Densification is achieved through 
a set of associated basis functions $\psi_k$:
%
\begin{equation}
u(\vec{x}) = \sum_k \psi_k(\vec{x}) \vec{u}_k
\label{eq:intp_kernel}
\end{equation}
%
In our implementation, $\psi_k$ is chosen to be a tensor-product B-Spline kernel
of degree 3 ($B_3$).
Then, introducing \eqref{eq:intp_kernel} into \eqref{eq:nodes_tfm} and replacing
$\psi$ by the actual kernel function, the transformation writes
%
\begin{equation}
\vec{c}_i' = \vec{c}_i + \sum_k \left[ \vec{u}_k \, \underset{n}{\prod} B_3( (\vec{c}_i - \vec{x}_k) \cdot \hat{\mathbf{e}}_d ) \right].
\label{eq:transformation}
\end{equation}

@@

Whereas related \glspl*{adf} introduced in \autoref{sec:methods_background}
make use of explicit level-set formulations to solve \eqref{eq:gradient_descent},
we alternatively use \emph{shape-gradients}
\cite{jehan-besson_dream2s:_2003,herbulot_segmentation_2006}.
Let $r(\vec{x})$ be an ``arbitrary'' function over the image domain,
and $\Omega$ a certain image region with $\Gamma_{l,m}$ its corresponding
outer boundary as defined in \autoref{sec:methods_map}.
We now derive the domain integral w.r.t. $t$:
\begin{equation*}
\frac{\partial}{\partial t} \int_\Omega r(\vec{x}') d\vec{x} = \int_\Omega \frac{\partial}{\partial t}r(\vec{x}') d\vec{x} - \int_{\Gamma_{l,m}} r(\vec{x}') \left\langle \frac{\partial \Gamma_{l,m} }{\partial t}, N_{\Gamma_{l,m}}\right\rangle d\vec{x}
\end{equation*}
where $\left\langle\frac{\partial\Gamma_{l,m}}{\partial t}, N_{\Gamma_{l,m}}\right\rangle$ is 
the projection of the boundary movement on the unit inward normal $N_{\Gamma_{l,m}}$. Assuming
that the region descriptors $\Theta_l$ vary slowly enough, we can consider
that $\frac{\partial}{\partial t} r(\vec{x}') = 0$ and thus:
\begin{equation}
\frac{\partial}{\partial t} \int_\Omega r(\vec{x}') d\vec{x} = 
- \int_{\Gamma_{l,m}} r(\vec{x}') \left\langle \frac{\partial \Gamma_{l,m} }{\partial t}, N_{\Gamma_{l,m}}\right\rangle d\vec{x}
\label{eq:shape_gradients}
\end{equation}

>>>>

% -*- root: 00-main.tex -*-
\section{Methods}
\label{sec:methods}
%
\subsection{Related work}
\label{sec:methods_background}

We suggest clustering the
current methodologies of template-based segmentation methods into three groups.
The first group typically adds a shape prior term to the energy functional of
\citep{chan_level_2005,chen_using_2002,cremers_kernel_2006,gastaud_combining_2004,wyatt_map_2003,paragios_level_2003,yezzi_variational_2003,gorthi_segmentation_2009,pohl_unifying_2005,pohl_bayesian_2006,wang_joint_2006,guyader_combined_2011,besson_dream2s_2003,bertero_illposed_1988,cote_tractometer_2013}
an evolving active contour \citep{bresson_variational_2006, chan_level_2005,
chen_using_2002,cremers_kernel_2006,gastaud_combining_2004}.
These methods generally have an explicit description of the expected relative boundary
locations of the object to be delineated, and some even model the statistical deviations
from this average shape. By including a coordinates mapping, it is possible to perform
active contours based registration between timesteps in a time-series or between different
images \citep{bertalmio_morphing_2000,wyatt_map_2003,paragios_level_2003,vemuri_joint_2003,
yezzi_variational_2003}.
A summary of these second set of techniques is performed in \citep{droske_mumfordshah_2009},
proposing two different approaches to applying the Mumford-Shah \citep{mumford_optimal_1989}
functional in joint registration and segmentation. Finally, a derivation of the
latter groups is composed by atlas-based segmentation methods
\citep{gorthi_segmentation_2009,gorthi_active_2011,pohl_unifying_2005,
pohl_bayesian_2006,wang_joint_2006}, where the prior
imposes consistent voxel-based classification of contiguous regions.
A comprehensive summary of the \gls{acwe}-derived methodologies with special attention
to joint registration-segmentation methods is found in \citep{gorthi_active_2011}.
\Gls{acwe} and level-set based formulations have been widely applied in image
processing, as reviewed in \citep{suri_shape_2002}.

@@

Even though the historical evolution
of joint segmentation and registration procedures is strongly linked
to the level sets approximation, our work is based in derivations of
the \gls{acwe} framework \citep{chan_active_2001}, and related to
\citep{guyader_combined_2011}.
{\color{red} {The main diverging points with respect to
\citep{gorthi_active_2011} are: 1) there is no need for an explicit level set function
$\Phi_G$, as we replace the level set gradient computation $N_{\Phi_G}$ with shape
gradients \citep{besson_dream2s_2003,herbulot_segmentation_2006};
2) regularization is also based on linear diffusion smoothing \citep{thirion_image_1998},
but we replace the Gaussian filtering by other constraints
studied in \citep{nagel_investigation_1986} to better the problem;
3) optimization is applied in the spectral
domain, observing anisotropic and inhomogeneous mappings along each direction.
With respect \citep{guyader_combined_2011}, the main differences are
the distance function, and the spectral solution to the optimization updates,
as we shall cover in \autoref{sec:numerical_implementation}.}}

@@

{\color{red} {Using the region descriptors derived in \autoref{sec:methods_map}, we propose
an \gls{adf}-like, \gls{acwe}-based, piece-wise constant, image segmentation
model (where the unknown is the deformation field)
\cite{chan_active_2001} with the energy functional obtained in
\eqref{eq:final_map_energy}. This inverse problem is ill-posed
\cite{bertero_illposed_1988,hadamard_sur_1902}.
In order to account for deformation field regularity and to render the
problem well-posed, we include limiting and regularization terms into
the energy functional \cite{morozov_linear_1975,tichonov_solution_1963}:
%
\begin{multline}
E(u) = \sum\limits_l \int_{\Omega'_l} \mdist{f}{l} \,d\vec{x} \\
+ \int \vec{u}^T \, A \, \vec{u} \, d\vec{x} + \int \tr\{(\nabla \vec{u}^T)^T B (\nabla \vec{u}^T)\} d\vec{x}
\label{eq:complete_energy}
\end{multline}
%
These regularity terms ensure that the segmenting contours in
\gls{dwi} space are still close to their native shape. The model
easily allows to incorporate inhomogeneous and anisotropic
regularization \cite{nagel_investigation_1986} to better regularize
the \gls{epi} distortion.}}

@@

\subsubsection{Deformation field}
\label{sec:deformation_field}
The transformation from ``T1-'' into ``\gls{dwi}-'' coordinate space is
achieved through a dense deformation field $u(\vec{x})$, such that:
\begin{equation}
\vec{c}_i' = U\{\vec{c}_i\} = \vec{c}_i + u(\vec{c}_i),
\label{eq:nodes_tfm}
\end{equation}
where $U$ is defined in \eqref{eq:deformation_field}. Since the nodes of the anatomical
surfaces might lay off-grid, it is required to derive $u(\vec{x})$ from a discrete
set of parameters $\{\vec{u}_k\}_{k=1 \ldots K}$. Densification is achieved through
a set of associated basis functions $\psi_k$:
%
\begin{equation}
u(\vec{x}) = \sum_k \psi_k(\vec{x}) \vec{u}_k
\label{eq:intp_kernel}
\end{equation}
%
In our implementation, $\psi_k$ is chosen to be a tensor-product B-Spline kernel
of degree 3 ($B_3$).
Then, introducing \eqref{eq:intp_kernel} into \eqref{eq:nodes_tfm} and replacing
$\psi$ by the actual kernel function, the transformation writes
%
\begin{equation}
\vec{c}_i' = \vec{c}_i + \sum_k \left[ \vec{u}_k \, \underset{d}{\prod} B_3( (\vec{c}_i - \vec{x}_k) \cdot \hat{\mathbf{e}}_d ) \right].
\label{eq:transformation}
\end{equation}

@@

Whereas related \glspl*{adf} introduced in \autoref{sec:methods_background}
make use of explicit level-set formulations to solve \eqref{eq:gradient_descent},
we alternatively use \emph{shape-gradients}
\cite{besson_dream2s_2003,herbulot_segmentation_2006}.
Let $r(\vec{x})$ be an ``arbitrary'' function over the image domain,
and $\Omega$ a certain image region with $\Gamma_{l,m}$ its corresponding
outer boundary as defined in \autoref{sec:methods_map}.
We now derive the domain integral w.r.t. $t$:
\begin{equation*}
\frac{\partial}{\partial t} \int_\Omega r(\vec{x}') d\vec{x} = \int_\Omega \frac{\partial}{\partial t}r(\vec{x}') d\vec{x} - \int_{\Gamma_{l,m}} r(\vec{x}') \left\langle \frac{\partial \Gamma_{l,m} }{\partial t}, N_{\Gamma_{l,m}}\right\rangle d\vec{x}
\end{equation*}
where $\left\langle\frac{\partial\Gamma_{l,m}}{\partial t}, N_{\Gamma_{l,m}}\right\rangle$ is
the projection of the boundary movement on the unit inward normal $N_{\Gamma_{l,m}}$. Assuming
that the region descriptors $\Theta_l$ vary slowly enough, we can consider
that $\frac{\partial}{\partial t} r(\vec{x}') = 0$ and thus:
\begin{equation}
\frac{\partial}{\partial t} \int_\Omega r(\vec{x}') d\vec{x} =
- \int_{\Gamma_{l,m}} r(\vec{x}') \left\langle \frac{\partial \Gamma_{l,m} }{\partial t}, N_{\Gamma_{l,m}}\right\rangle d\vec{x}
\label{eq:shape_gradients}
\end{equation}

### 2014-10-20 Publications/2013-TMI/05-experiments.tex
update TMI paper

\section{Data and experiments}
\label{sec:experiments}
%

@@

\subsubsection{Simulated digital phantom} %
%
The lack of a widely accepted \emph{gold}-standard in the application
field has been addressed by several authors \citep{cote_tractometer:_2013}
In this work, we use one of the most complete, and publicly available,
digital phantoms (\url{http://hardi.epfl.ch/static/events/2013_ISBI/testing_data.html}).
The phantom is a spherical volume containing a set of fiber bundles, that connect 
one area of a ``cortex'' to another. The model accounts for \gls{pve} using 
a similar approach to \citep{close_software_2009} and for \gls{csf} contamination
as well.

>>>>

% -*- root: 00-main.tex -*-
\section{Data and experiments}
\label{sec:experiments}
%

@@

\subsubsection{Simulated digital phantom} %
%
The lack of a widely accepted \emph{gold}-standard in the application
field has been addressed by several authors \citep{cote_tractometer_2013}
In this work, we use one of the most complete, and publicly available,
digital phantoms (\url{http://hardi.epfl.ch/static/events/2013_ISBI/testing_data.html}).
The phantom is a spherical volume containing a set of fiber bundles, that connect
one area of a ``cortex'' to another. The model accounts for \gls{pve} using
a similar approach to \citep{close_software_2009} and for \gls{csf} contamination
as well.
